Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-30 mu/article/prb/article.tex
First manuscript sent to PRB

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials is short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimensional (2D)
electronic systems; however, numerous important electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}. 
In this paper, we offer two 50\% hydrogenated graphene
noncentrosymmetric structures, both demonstrating a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

@@

We reported the results of an \emph{ab initio} calculations for the spin
velocity injection (SVI) due to the one-photon absorption of the linearly
polarized light in the Up and Alt 2D 50\% hydrogenated graphene structures.
Different possible arrangements of the of the spin injection have been
considered: we made the calculations for the cases when the spin is polarized
in $z$ direction or when the velocity is directed along $x$ or $y$. To the best
of our knowledge, this effect  has not been previously reported in this  2D
partially hydrogenated structures. We have shown that the  SVI demonstrates an
anisotropic behavior, which is very sensitive to the symmetry of the structures
of interest. We have found that the Up structure shows the strongest response
for the spin directed along $z$, resulting in the velocity
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 668.0$\,Km/s for the
incoming photon energy of 0.084\,eV. Also, the Alt structure has the strongest
response when the spin moves along $y$ direction, resulting in
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s for the incoming
photon energy of 0.720\,eV. The speed values obtained here are of the same
order of magnitude as those of Ref.~\onlinecite{najmaiePRB03} in unbiased
semiconductor quantum well structures, while they are of order of magnitude
higher compared to 3D bulk materials. The spin relaxation time in pure and
doped graphene goes  from nanoseconds to milliseconds,
\cite{wojtaszekPRB13,ertlerPRB09} that, according to our results, 
is long enough in view of the
high spin velocity transport that we obtain for both
structures. Therefore, the Up and the Alt graphene structures
considered here are  excellent
candidates for the development of spintronics devices that require pure spin
current (PSC).

\section{Acknowledgment}
This work has been supported by \emph{Consejo Nacional de Ciencia y
Tecnolog\'ia} (CONACyT), M\'exico, Grant No. 153930.
R.Z.P. thanks CONACyT for scholarship support.

>>>>

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in semiconducting materials can be too short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimensional (2D)
electronic systems; however, numerous important electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}.

In this paper, we offer two 50\% hydrogenated graphene
noncentrosymmetric structures, both demonstrating a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

@@

We reported the results of an \emph{ab initio} calculations for the spin
velocity injection (SVI) due to the one-photon absorption of the linearly
polarized light in the Up and Alt 2D 50\% hydrogenated graphene structures.
Different possible arrangements of the of the spin injection have been
considered: we made the calculations for the cases when the spin is polarized
in $z$ direction or when the velocity is directed along $x$ or $y$. To the best
of our knowledge, this effect has not been previously reported in these 2D
partially hydrogenated structures. We have shown that the  SVI demonstrates an
anisotropic behavior, which is very sensitive to the symmetry of the structures
of interest. We have found that the Up structure shows the strongest response
for the spin directed along $z$, resulting in the velocity
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 668.0$\,Km/s for the
incoming photon energy of 0.084\,eV. Also, the Alt structure has the strongest
response when the spin moves along $y$ direction, resulting in
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s for the incoming
photon energy of 0.720\,eV. The speed values obtained here are of the same
order of magnitude as those of Ref.~\onlinecite{najmaiePRB03} in unbiased
semiconductor quantum well structures, while they are of order of magnitude
higher compared to 3D bulk materials. Considering the fact that the spin
relaxation time in pure and doped graphene ranges from nanoseconds to
milliseconds, \cite{wojtaszekPRB13,ertlerPRB09} and, in view of the high spin
velocity transport that we obtained for the both structures, this time is
sufficiently long enough to have the SVI effect observed experimentally.
Therefore, the Up and the Alt graphene structures considered here are
excellent candidates for the development of spintronics devices that require
pure spin current (PSC).

\section{Acknowledgment}
This work has been supported by \emph{Consejo Nacional de Ciencia y
Tecnolog\'ia} (CONACyT), M\'exico, Grant No. 153930.
A.I.S thanks to Centro de Investigaciones en Optica (CIO) for the hospitality 
during his sabbatical research leave.
R.Z.P. thanks CONACyT an CIO for scholarship support.
